Energy Dependent X-Ray Rim Widths in Tycho's Supernova Remnant

### 2014-09-10 manuscript/tycho-paper.tex
Tablepalooza paper draft, fit avg Tycho fwhms
Parse lots of manual output (code coming soon) to add tables to draft
Set up Tycho fitting ipynb to work on average FWHMs

\documentclass[manuscript]{aastex}  % one-column, double-spaced
% \documentclass[12pt,preprint]{aastex}  % one-column, single-spaced
% \documentclass[iop, apj, twocolappendix]{emulateapj}

@@

\title{Energy Dependent X-Ray Rim Width in Tycho's Supernova Remnant}

@@

Like SN 1006, Tycho has well defined synchrotron rims associated with expansion
into a low-density ISM \note[Aaron]{justification: high density ISM would
prevent the forward shock from running ahead of the thermal ejecta?}.
\citet{williams2013} favor mean pre-shock ISM density $\abt 0.2 \unit{cm^{-3}}$
from \textit{Spitzer} observations of shocked ISM dust emission,
\note[Aaron]{is this proton or electron density?  Or are they equal pre-shock?
Text mentions $n_e = 1.2 n_p$ \emph{post}-shock.}
consistent with Tycho's X-ray expansion \citep{katsuda2010}.
Compare w/ SN 1006 pre-shock density $n_p \abt 0.15$--$0.3 \unit{cm^{-3}}$
\citep{raymond2007, heng2007, winkler2013}.

@@

Any other remarks? (comparison to core-collapse SNe? no compact object to mess
things up? or is that flatly trivial)

@@

\note[Aaron]{EXPLAIN what the equations mean.  How do we eventually back out a filament
shape, filament width?  Just a few sentences.}

Bibdesk -- mark up all the papers I've really gone through and understood.
Then, add the relevant citations and notes, to be sure I haven't missed or
misunderstood anything.

@@

% TODO go through and double check tenses.

We measured synchrotron rim widths from an archival \Chandra
ACIS-I observation of Tycho
(RA: 00\tsup{h}25\tsup{m}19\fs0, dec: +64\arcdeg08\arcmin10\farcs0; J2000)
between 2009 Apr 11 and 2009 May 5 (PI: John P. Hughes;
\dataset[ADS/Sa.CXO\#obs/10093--10097]{ObsIDs: 10093--10097},
\dataset[ADS/Sa.CXO\#obs/10902--10906]{10902--10906}).
The total exposure time is $734 \unit{ks}$.
Level 1 \Chandra data were reprocessed with CIAO 4.6 and CALDB 4.6.1.1 and kept
unbinned with ACIS spatial resolution $0.492\arcsec$.
Merged and corrected events were divided into five energy bands:
0.7--1 keV, 1--1.7 keV, 2--3 keV, 3--4.5 keV, and 4.5--7 keV;
we excluded 1.7--2 keV counts to avoid \ion{Si}{13} (He$\alpha$) line
emission, prevalent in the remnant's thermal ejecta, which could contaminate
our nonthermal profile measurements.

@@

\begin{figure}
    \plotone{figures/f0-snr.eps}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).
    \note[Aaron]{Temporary figure; regions need re-numbering.}}
    \label{fig:snr}
\end{figure}

@@

To estimate FWHM uncertainties, we horizontally stretched each best-fit
profile by mapping radial coordinate $x$ to
$x'(x) = x (1 + \xi (x-x_0)/(50\arcsec-x))$ with $\xi$ an arbitrary stretching
parameter; this yields a new profile $h'(x) = h(x'(x))$.
We varied $\xi$ (and hence rim FWHM) to vary each profile fit $\chi^2$ by 2.7,
yielding confidence limits on our measured FWHMs

@@

As a ``back-verification'', we check that our rims' X-ray spectra
are clear of thermal line emission.

In each radial profile, we select an interval containing FWHM bounds from all
energy bands.  We

Each rim is divided into an upstream

From our model fits and we divide the rim into an upstream and downstream region;
the

And we shall purify the remnant's rims\\
that they may show us how they do thin\\
discriminating damped models.

This also gives a post-hoc / further filter on intermediate results --
we can quantify the thermal emission...

We further constrain thermal emission by checking that spectral fits
favor pure continuum emission over continuum with Gaussian components for
\ion{Si}{13} and \ion{S}{15} He$\alpha$ lines at $\abt 1.8$, $2.4$ keV.

HERE I should mention any upper bounds on spectral lines (Equivalent width or
what have you).  Since this is not a ``result'', being ancillary to our work.

@@

@@

To perform fits, we tabulated a large grid for fixed values of $\mu$ and shock
velocity $v_s$.

If any fit runs to $\eta_2 = 10^5$ or $B_0 = 10 \unit{mG}$ (for Tycho),
we deem it effectively unconstrained.

@@

We present results here.  Explain a bit about how we divvy up / present the
results.  Also, for each region, give the relevant shock velocity value.  Or, a
range of values.
We do X to the FWHMs (geometric/arithmetic average) within each filament.
We also consider global FWHMs average.

% --------------------
% FWHM results, tables
% --------------------
\subsection{Measured FWHMs, energy dependence, etc}

\begin{table*}
\scriptsize
\centering
\caption{Filament 2 full width at half max measurements.
Exponent $m_E$ estimated point to point.\label{tab:flmt2}}
\input{tab-fwhms.tex}
\tablecomments{Mean FWHMs computed by arithmetic mean, errors from blah}
\end{table*}

@@

% -------------------------
% Model fit results, tables
% -------------------------
\subsection{Model}

Table reference goes here (Table~\ref{tab:fit1}).
Explain error values on parameters.

\begin{table}
\centering
\caption{Full model best fit parameters, region 1.\label{tab:fit1}}
\input{tab-fits.tex}
\end{table}

@@

What do all these numbers with their errors mean?

@@

Diffusion -- sub-Bohm diffusion?

\subsection{Comparisons to Cas A (pulled from Sean's \LaTeX\ source)}

\note[Aaron]{for my own reference/reading/thinking (temporary)}

To quickly compare how our results will differ from those of other authors, we
applied the analytic approximation of Equation 22 and Equation 9 to the FWHMs
presented by Araya et al. (2010) for Cas A. Here we used $m_{\rm E} = \log(
{\rm FWHM}(3keV)/{\rm FWHM}(.3keV))/\log(10)$ and fit the filaments at a photon
energy of 3 keV.  Hence $\eta_3 \equiv D/D_{\rm Bohm}(3 keV)$. Using the same
parameters, our results are shown in Table~\ref{tab:Araya}. To fit the energy
dependence of the FWHM (i.e. $m_{\rm E}$), we required higher diffusion
coefficients and, consequently, magnetic fields about an order of magnitude
higher than their results. We note here that values of $\mu<1$ were unable to
reproduce the data and that the diffusion coefficients cited by Araya et al.
(2010) are below the minimally allowed Bohm value ($\eta_3 = 1$).

\begin{table*}[h]
\caption{Best fit parameters for the filaments of Cas A based on data from
Araya et. al (2010) in varying values of $\mu$ (Approximate Analytic Results)
Dashes denote places where fits were unobtainable.}
\centering
\begin{tabular}{c c c c c c c c c}
\hline \hline
  &\multicolumn{2}{c}{$\mu = 1$} & \multicolumn{2}{c}{$\mu = 1.5$} & \multicolumn{2}{c}{$\mu = 2$} & \multicolumn{2}{c}{Araya et. al} \\
  & & & & & &  &\multicolumn{2}{c}{($\mu = 1$)} \\
$Filament$ &$\eta_{3}$ & $B_{0}$ &$\eta_{3}$ & $B_{0}$ & $\eta_{3}$ & $B_{0}$& $\eta_{3}$ & $B_{0}$ \\ [.5ex]
1 & 6.5  & 750 $\mu$G & 2 & 567 $\mu$G & 1.1 & 506 $\mu$G & 0.12 & 72 $\mu$G \\
2 & 3.6 & 710 $\mu$G & 1.4 & 582 $\mu$G & - & - & 0.02 & 37 $\mu$G\\
3 & 3.2 &  710 $\mu$G& 1.3 & 589 $\mu$G& - & - & 0.02 & 53 $\mu$G\\
4 & 3.1& 515 $\mu$G & 1.3 & 430 $\mu$G & -& - & 0.02 & 40 $\mu$G \\
5 & 10.7 & 1163 $\mu$G & 2.5 & 809 $\mu$G & 1.3&  706 $\mu$G & 0.025 & 52 $\mu$G \\
6 & 5.7 & 844 $\mu$G & 1.9 & 650 $\mu$G  & 1 & 583 $\mu$G & 0.1 & 56 $\mu$G \\
7 & 8.3 & 872 $\mu$G & 2.2 & 635 $\mu$G& 1.2 & 560 $\mu$G & 0.15 & 66 $\mu$G\\
8 & 13 & 1010 $\mu$G & 2.7 & 738 $\mu$G & 1.4 & 639 $\mu$G & 0.02 & 35 $\mu$G\\
9 & 4.7 & 605 $\mu$G & 1.7 & 479 $\mu$G & - & -& 0.02 & 29 $\mu$G\\

\hline
\hline
\end{tabular}
\label{tab:Araya}
\end{table*}

\subsection{End of Sean's Cas A analysis}

Lalala.

@@

@@

For comparison to \citetalias{ressler2014}, we present our full model results
employing 2 energy bands (as in \citetalias{ressler2014}) and all 3 energy
bands (0.7--1, 1--2, 2--7 keV).  The simple model results are identical,
although the errors obtained from my procedure are larger.

\section{SN 1006 results, 2 energy bands}  % Use numberedappendix option for emulateapj

Tables tables tables tables

\section{SN 1006 results, 3 energy bands}

Tables tables tables tables

\section{Discussion}

Lalalalala.  Beautiful numbers!

>>>>

% \documentclass[manuscript]{aastex}  % one-column, double-spaced
% \documentclass[12pt,preprint]{aastex}  % one-column, single-spaced
\documentclass[iop, apj, numberedappendix, twocolappendix]{emulateapj}
% \documentclass[iop, apj, numberedappendix]{emulateapj}

@@

\title{Energy Dependent X-Ray Rim Widths in Tycho's Supernova Remnant}

@@

Like SN 1006, Tycho has well defined synchrotron rims associated with expansion
into a low-density ISM \note[Aaron]{justification: high density ISM would
prevent the forward shock from running ahead of the thermal ejecta?}.
\citet{williams2013} favor mean pre-shock ISM density $\abt 0.2 \unit{cm^{-3}}$
from \textit{Spitzer} observations of shocked ISM dust emission,
\note[Aaron]{is this proton or electron density?  Or are they equal pre-shock?
Text mentions $n_e = 1.2 n_p$ \emph{post}-shock.}
consistent with Tycho's X-ray expansion \citep{katsuda2010}.
Compare w/ SN 1006 pre-shock density $n_p \abt 0.15$--$0.3 \unit{cm^{-3}}$
\citep{raymond2007, heng2007, winkler2013}.  \note[Aaron]{ISM gradients?}

@@

Previous estimates of Tycho fields/diffusion (something similar to Sean's table
for SN 1006?).  CR acceleration.

@@

Explain how the transport models constrain diffusion via $m_E$, define $m_E$.
Explain how we back out a filament shape/width from these models

\subsection{Magnetic damping}

We also consider a magnetically damped field of form
\[
    B(x) = B_{\mt{min}} + \left(B_0 + B_{\mt{min}}\right) \exp\left(-x / a_b\right) ,
\]
following \citetalias[Section 3.2]{ressler2014}.

@@

% TODO go through and double check tenses.
\note[Aaron]{inconsistent present/past tense usage}
We measured synchrotron rim widths from an archival \Chandra
ACIS-I observation of Tycho
(RA: 00\tsup{h}25\tsup{m}19\fs0, dec: +64\arcdeg08\arcmin10\farcs0; J2000)
between 2009 Apr 11 and 2009 May 5 (PI: John P. Hughes;
\dataset[ADS/Sa.CXO\#obs/10093--10097]{ObsIDs: 10093--10097},
\dataset[ADS/Sa.CXO\#obs/10902--10906]{10902--10906}).
The total exposure time was $734 \unit{ks}$.
Level 1 \Chandra data were reprocessed with CIAO 4.6 and CALDB 4.6.1.1 and kept
unbinned with ACIS spatial resolution $0.492\arcsec$.
Merged and corrected events were divided into five energy bands:
0.7--1 keV, 1--1.7 keV, 2--3 keV, 3--4.5 keV, and 4.5--7 keV;
we excluded 1.7--2 keV counts to avoid \ion{Si}{13} (He$\alpha$) line
emission, prevalent in the remnant's thermal ejecta, which could contaminate
our nonthermal profile measurements.

@@

\begin{figure}
    %\plotone{figures/f0-snr.pdf}
    \centering
    \includegraphics[width=0.35\textwidth]{figures/f0-snr.pdf}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).
    \note[Aaron]{Temporary figure; regions need re-numbering.}}
    \label{fig:snr}
\end{figure}

@@

To estimate FWHM uncertainties, we horizontally stretched each best-fit
profile by mapping radial coordinate $x$ to
$x'(x) = x (1 + \xi (x-x_0)/(50\arcsec-x))$ with $\xi$ an arbitrary stretching
parameter; this yields a new profile $h'(x) = h(x'(x))$.
We varied $\xi$ (and hence rim FWHM) to vary each profile fit $\chi^2$ by 2.7,
yielding confidence limits on our measured FWHMs.

@@

We extracted and fitted spectra for all regions to confirm that our rim width
measurements are not contaminated by thermal line emission.
We divided each region into an upstream and downstream band.  The upstream band
is the smallest sub-region containing FWHM bounds from all energy bands.
The downstream band extends from the back of the rim (where the upstream band
ends) to the rim model fit's (equation~\eqref{eq:prof}) downstream limit.
Figure~\ref{fig:spec} plots an example profile (Region 1, 4.5--7 keV) with the
downstream/upstream selections highlighted in light blue/gray respectively.
We fitted spectra from each region to an absorbed power law between $0.5$ and $7$
keV in XSPEC.

TODO?! We also fit the spectra to absorbed power law models with 1 and 2
additional Gaussian components for \ion{Si}{13} and \ion{S}{15} He$\alpha$
lines at $\abt 1.8$, $2.4$ keV.  We compute some equivalent width, or compare
the goodness-of-fit with and without the additional lines, to rule out the
presence of a line (but, cf. \citet{protassov2002} on line detection
statistics!).

@@

We fitted the two transport models given by Equations~\eqref{eq:simp-mod} and
\eqref{eq:full-mod} to the measured data.  Here we present only the full
model (equation~\eqref{eq:full-mod}) fits.  \note[Aaron]{Could we present the
full model alone, given that the numbers / qualitative trends for both models
are fairly similar?  In \citetalias{ressler2014} Sean used the simple model to
build intuition and gain physical insight, to great effect.  But, will we need
that here?}

@@

To perform fits, we tabulated a large grid for fixed values of $\mu$ and shock
velocity $v_s$.  If any fit runs to $\eta_2 = 10^5$ or $B_0 = 10 \unit{mG}$,
we deem it effectively unconstrained (only a few fits do so, and only to the
$\eta_2$ limit).
Note that we do not allow / consider fits with $\eta_2$ negative (may be
trivial, just note parameter bounds).

@@

Explain a bit about how we divvy up / present the results.  What values are we
averaging, where do our errors come from?
For each region, give the relevant shock velocity value (and a range of
values?).
We do X to the FWHMs (geometric/arithmetic average) within each filament.

% --------------------
% FWHM results, tables
% --------------------
\subsection{Rim fits, spectra, and width measurements}

We give figures and tables with rim profile fit information.

\begin{table*}[h]
\scriptsize
\centering
\caption{Filament 2 full width at half max measurements.
Exponent $m_E$ estimated point to point.\label{tab:flmt2}}
\input{tab-fwhms.tex}
\tablecomments{Mean FWHMs computed by arithmetic mean, errors from blah}
\end{table*}

@@

Here, mention any upper bounds on spectral lines (equivalent width or w/e).
Since this is not a ``result'', being ancillary to our work.

Table of best fit chi-squares, power law indices for spectral fits...

% -------------------------
% Model fit results, tables
% -------------------------
\subsection{Model fit results}

Here are some fit results for a subset of regions, just those contained in
Filament 1 (Table~\ref{tab:fit}).  Errors are estimated by 
Explain error values on parameters

\begin{table*}
\scriptsize
\centering
\caption{Full model best fit parameters, Filament 1.
    \label{tab:fit}}
\input{tab-fits-flmt1.tex}
\end{table*}

Table~\ref{tab:fit-avg-fwhm} shows best fits to averaged FWHMs (averaged FWHMs
not shown).  For FWHM errors, I use the standard error of the mean (with some
minor nuances; I am using sample standard deviation).

\begin{table*}
\scriptsize
\centering
\caption{Full model best fit parameters for filament-averaged FWHMs.
    \label{tab:fit-avg-fwhm}}
\input{tab-fits-avg.tex}
\end{table*}

Finally, Table~\ref{tab:fit-avg-par} shows averaged best-fit parameters for
Filament 1 only, at this time.  The parameter errors are standard errors from
the averaging process and may reflect the presence of outliers.

\begin{table*}
\scriptsize
\centering
\caption{Average of best-fit parameters for individual regions
    \label{tab:fit-avg-par}}
\input{tab-fits-par-avg.tex}
\end{table*}

\begin{table*}
\scriptsize
\centering
\caption{Full model best fit parameters, Filament ??, with $\eta_2 = 1$ fixed}
%    \label{tab:fit-avg-par}}
%\input{tab-fits-par-avg.tex}
Coming soon...
\end{table*}

@@

What do all these numbers with their errors mean?
Can we favor a value of $\mu$ (and hence a turbulent energy spectrum)?

How meaningful are our parameter estimates and errors?

@@

Diffusion -- sub-Bohm diffusion in Filament 4?

Constraints on diffusion / B field, if any.  What impact this has for models of
B amplification etc.

@@

\clearpage

@@

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

% --------------
% SN 1006 tables
% --------------
\section{Full model validation, SN 1006 (DRAFT ONLY)}

For comparison to \citetalias{ressler2014}, we reproduce Sean's table
(Table~\ref{tab:sean} and present full model fits with 2 and 3 energy bands
(Tables~\ref{tab:sn1006-2band} and \ref{tab:sn1006-3band}, respectively).
We don't compare simple model results, as they are identical (only the error
calculations differ).
The procedure of Tables~\ref{tab:sean},~\ref{tab:sn1006-2band} are not quite
the same.  Sean used $m_E$ at 2 keV instead of the $1$--$2$ keV width for
fitting, whereas I used both widths.  But, it should not matter much as
the fit often has $\chi^2 \sim 0$, either way (depending on the exact
measurements and values of $\mu$).

Please take the errors with a grain of salt.  They are automatically
generated and \textbf{have not been manually validated}.  Some values may be
invalid, where my error-finding code failed and gave a best, conservative guess
of the error.

\begin{table}[h]
\scriptsize
\centering
\caption{Sean's SN 1006 best fit parameters \citepalias[Table 8]{ressler2014}}
\label{tab:sean}
\input{tab-sean.tex}
\end{table}

\begin{table*}[h]
\scriptsize
\centering
\caption{SN 1006 best fit parameters, 2 highest energy bands (full model)}
\label{tab:sn1006-2band}
\input{tab-sn1006-2band.tex}
\end{table*}

\begin{table*}[h]
\scriptsize
\centering
\caption{SN 1006 best fit parameters, 3 energy bands (full model)}
\label{tab:sn1006-3band}
\input{tab-sn1006-3band.tex}
\end{table*}

% ------------------
% Sean's Cas A table
% ------------------
\section{Sean's Cas A comparison (DRAFT ONLY)}

This is dredged from the \LaTeX source for Sean's paper, available on arXiv.
I put this here just for my own reference.

To quickly compare how our results will differ from those of other authors, we
applied the analytic approximation of Equation 22 and Equation 9 to the FWHMs
presented by Araya et al. (2010) for Cas A. Here we used $m_{\rm E} = \log(
{\rm FWHM}(3keV)/{\rm FWHM}(.3keV))/\log(10)$ and fit the filaments at a photon
energy of 3 keV.  Hence $\eta_3 \equiv D/D_{\rm Bohm}(3 keV)$. Using the same
parameters, our results are shown in Table~\ref{tab:araya}. To fit the energy
dependence of the FWHM (i.e. $m_{\rm E}$), we required higher diffusion
coefficients and, consequently, magnetic fields about an order of magnitude
higher than their results. We note here that values of $\mu<1$ were unable to
reproduce the data and that the diffusion coefficients cited by Araya et al.
(2010) are below the minimally allowed Bohm value ($\eta_3 = 1$).

\begin{table}[h]
\tiny
\centering
\caption{Best fit parameters for the filaments of Cas A based on data from
Araya et. al (2010) in varying values of $\mu$ (Approximate Analytic Results)
Dashes denote places where fits were unobtainable.}
\label{tab:araya}
\input{tab-sean-casA.tex}
\end{table}
